---
title: "Hw7"
author: "Jaxon Lara"
date: "jtl3236"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(mosaic)
#install.packages("MatchIt")
library(MatchIt)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Numbers and statistics
armfold = read.csv("armfold.csv")

bootstrap_diffPropMF = do(5000)*(
  prop(resample(armfold[armfold$Sex == "Male",]$LonR_fold == 1)) -
  prop(resample(armfold[armfold$Sex == "Female",]$LonR_fold == 1))
)
diffPropMF = confint(bootstrap_diffPropMF, level=0.95)

numMales = sum(armfold$Sex=="Male")
numFemales = numFemale = sum(armfold$Sex=="Female")

propMalesLeft = as.numeric(prop(armfold[armfold$Sex == "Male",]$LonR_fold == 1))
propFemalesLeft = as.numeric(prop(armfold[armfold$Sex == "Female",]$LonR_fold == 1))

standard_error = sqrt( ((propMalesLeft*(1 - propMalesLeft)) / numMales) + ((propFemalesLeft*(1 - propFemalesLeft)) / numFemales) )

mean_sampling_distribution = mean(bootstrap_diffPropMF$prop_TRUE)

lowerbound = mean_sampling_distribution - 1.96 * standard_error
upperbound = mean_sampling_distribution + 1.96 * standard_error
```

# 1A  
The number of male students: `r numMales`  
The number of female students: `r numFemales`  
The sample proportion of males who folded their left arm on top: `r round(propMalesLeft,2)`  
The sample proportion of males who folded their left arm on top: `r round(propFemalesLeft,2)`  
  
# 1B  
The observed difference in proportions between the two groups (males - females): `r propMalesLeft - propFemalesLeft`  

# 1C  
The 95% confidence interval for the difference in proportions (males - females). (R's built in function)  
Lower bound: `r round(diffPropMF$lower,3)`, Upper bound: `r round(diffPropMF$upper,3)`  

standard error for the difference in proportions:  
sqrt( ((samplePropGroup1(1-samplePropGroup1)) / sampleSize1) + ((samplePropGroup2(1-samplePropGroup2)) / sampleSize2)  
or  
sqrt( ((propMalesLeft(1-propMalesLeft)) / numMales) + ((propFemalesLeft(1-propFemalesLeft)) / numFemales) )  

standard error for difference in proportions: `r round(standard_error,3)`  

sample mean for the difference in proportions: `r round(mean_sampling_distribution,3)`  

z, standard deviations: 1.96  

upperbound = (sample mean) + (1.96) * (standard error)  
lowerbound = (sample mean) - (1.96) * (standard error)  
upper bound: `r upperbound`  
lower bound: `r lowerbound`  

# 1D  
If we were to take many many samples of students who were asked to fold their arms across their chest  
and calculate their corresponding confidence intervals. Then we would expect that 95% of those confidence  
intervals contain the true difference in proportions.  

# 1E  
The standard error for the difference in proportions represents how varied a sample mean for the  
difference in proportions could be from the true population difference in proportions.  

# 1F  
The term sampling distribution refers to the distribution of bootstrapped samples created from resampling  
the proportion of Males/Females who folded their left arm on top. Specifically, isolating the cases whose  
sex was male (fixed sex) and resampling whether or not they folded their left arm on top with replacement  
(varies from sample to sample). Then doing the same for cases whose sex was female and subtracting the  
male proportion from the female proportion to get that sample's difference in proportions. Then repeat.  

# 1G  
The central limit theorem, because the sampling distribution is reasonably assumed to be made up  
of many many samples. The central limit theorem says that the distribution of a sample mean or in this  
case proportions will become approximately normal with a reasonably large sample size.  

# 1H  
I would say that their assessment is correct because the 95% confidence interval includes 0  
and so there is not enough evidence to suggest there is a difference in arm folding based on sex.  

# 1I  
Yes the confidence intervals will be different across the samples but the true difference in  
proportions of folding left arms on top for males and females should be contained within  
95% of those confidence intervals.  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
turnout = read.csv("turnout.csv")

#Prop of those receiving a GOTV call who voted in 1998
call_1998 = prop(turnout[turnout$GOTV_call == 1,]$voted1998 == 1)

#Sample proportion of those not receiving a GOTV call who voted in 1998
notCall_1998 = prop(turnout[turnout$GOTV_call == 0,]$voted1998 == 1)

#95% confidence interval by bootstrapping
bootstrap_diffPropTO = do(5000)*(
  prop(resample(turnout[turnout$GOTV_call == 1,]$voted1998 == 1)) -
  prop(resample(turnout[turnout$GOTV_call == 0,]$voted1998 == 1))
)
confintTO = confint(bootstrap_diffPropTO, level=0.95)
```

# 2A  
The proportion of those receiving a GOTV call who voted in 1998: `r round(call_1998,2)`  
The sample proportion of those not reciving a GOTV call who voted in 1998: `r round(notCall_1998,2)`  

Large-sample 95% confidence interval for the difference in these two proportions:  
Lower bound: `r round(confintTO$lower,3)`, Upper bound: `r round(confintTO$upper,3)`  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#voted1996
bootstrap_voted1996_1 = do(5000)*cor(voted1996 ~ voted1998, data=resample(turnout)) #outcome
confint1996_1 = confint(bootstrap_voted1996_1, level=0.95)

bootstrap_voted1996_2 = do(5000)*cor(voted1996 ~ GOTV_call, data=resample(turnout)) #treatment
confint1996_2 = confint(bootstrap_voted1996_2, level=0.95)

#AGE
bootstrap_AGE_1 = do(5000)*cor(AGE ~ voted1998, data=resample(turnout)) #outcome
confintAGE_1 = confint(bootstrap_AGE_1, level=0.95)

bootstrap_AGE_2 = do(5000)*cor(AGE ~ GOTV_call, data=resample(turnout)) #treatment
confintAGE_2 = confint(bootstrap_AGE_2, level=0.95)

#MAJORPTY
bootstrap_MAJORPTY_1 = do(5000)*cor(MAJORPTY ~ voted1998, data=resample(turnout)) #outcome
confintMAJORPTY_1 = confint(bootstrap_MAJORPTY_1, level=0.95)

bootstrap_MAJORPTY_2 = do(5000)*cor(MAJORPTY ~ GOTV_call, data=resample(turnout)) #treatment
confintMAJORPTY_2 = confint(bootstrap_MAJORPTY_2, level=0.95)
```

# 2B
The correlation between voted1996 and voted1998 using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confint1996_1$lower,3)`, Upper bound: `r round(confint1996_1$upper,3)`  
The correlation between voted1996 and GOTV_call call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confint1996_2$lower,3)`, Upper bound: `r round(confint1996_2$upper,3)`  
voted1996 is highly correlated to both the treatment(GOTV_call) and the outcome(voted1998)  
because 0 is included in neither interval which is evidence that voted1996 is a confounder.  

The correlation between AGE and voted1998 using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintAGE_1$lower,3)`, Upper bound: `r round(confintAGE_1$upper,3)`  
The correlation between AGE and GOTV_call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintAGE_2$lower,3)`, Upper bound: `r round(confintAGE_2$upper,3)`  
AGE is highly correlated to both the treatment(GOTV_call) and the outcome(voted1998)  
because 0 is included in neither interval which is evidence that AGE is a confounder.  

The correlation between MAJORPTY and voted1998 using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintMAJORPTY_1$lower,3)`, Upper bound: `r round(confintMAJORPTY_1$upper,3)`  
The correlation between MAJORPTY and GOTV_call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintMAJORPTY_2$lower,3)`, Upper bound: `r round(confintMAJORPTY_2$upper,3)`  
MAJORPTY is highly correlated to both the treatment(GOTV_call) and the outcome(voted1998)
because 0 is included in neither interval which is evidence that MAJORPTY is a confounder.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Match to the 3 confounders, voted1996, age, majorpty
turnout_match = matchit(GOTV_call ~ voted1996 + factor(AGE) + factor(MAJORPTY), data = turnout, ratio=5)
# Extracting the matched pairs
turnout_matched = match.data(turnout_match)



#voted1996
bootstrap_voted1996_2_matched = do(5000)*cor(voted1996 ~ GOTV_call, data=resample(turnout_matched)) #treatment
confint1996_3 = confint(bootstrap_voted1996_2_matched, level=0.95)

#AGE
bootstrap_AGE_2_matched = do(5000)*cor(AGE ~ GOTV_call, data=resample(turnout_matched)) #treatment
confintAGE_3 = confint(bootstrap_AGE_2_matched, level=0.95)

#MAJORPTY
bootstrap_MAJORPTY_2_matched = do(5000)*cor(MAJORPTY ~ GOTV_call, data=resample(turnout_matched)) #treatment
confintMAJORPTY_3 = confint(bootstrap_MAJORPTY_2_matched, level=0.95)



#Prop of those receiving a GOTV call who voted in 1998
call_1998_matched = prop(turnout_matched[turnout_matched$GOTV_call == 1,]$voted1998 == 1)

#Sample proportion of those not receiving a GOTV call who voted in 1998
notCall_1998_matched = prop(turnout_matched[turnout_matched$GOTV_call == 0,]$voted1998 == 1)

#95% confidence interval by bootstrapping (for matched dataset)
bootstrap_diffPropTO_matched = do(5000)*(
  prop(resample(turnout_matched[turnout_matched$GOTV_call == 1,]$voted1998 == 1)) -
  prop(resample(turnout_matched[turnout_matched$GOTV_call == 0,]$voted1998 == 1))
)
confintTO_matched = confint(bootstrap_diffPropTO_matched, level=0.95)
```

# 2C  
The correlation between voted1996 and GOTV_call call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confint1996_3$lower,3)`, Upper bound: `r round(confint1996_3$upper,3)`  
voted1996 is no longer highly correlated to the treatment(GOTV_call) because 0 is included  
in the confidence interval which is evidence that voted1996 is no longer a confounder.  

The correlation between AGE and GOTV_call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintAGE_3$lower,3)`, Upper bound: `r round(confintAGE_3$upper,3)`  
AGE is no longer highly correlated to the treatment(GOTV_call) because 0 is included  
in the confidence interval which is evidence that AGE is no longer a confounder.  

The correlation between MAJORPTY and GOTV_call using a large-sample confidence interval  
is estimated to be the following.  
Lower bound: `r round(confintMAJORPTY_3$lower,3)`, Upper bound: `r round(confintMAJORPTY_3$upper,3)`  
MAJORPTY is no longer highly correlated to the treatment(GOTV_call) because 0 is included  
in the confidence interval which is evidence that MAJORPTY is no longer a confounder.  


Repeating the analysis from part A with the matched data:  
The proportion of those receiving a GOTV call who voted in 1998: `r round(call_1998_matched,2)`  
The sample proportion of those not reciving a GOTV call who voted in 1998: `r round(notCall_1998_matched,2)`  

Large-sample 95% confidence interval for the difference in these two proportions:  
Lower bound: `r round(confintTO_matched$lower,3)`, Upper bound: `r round(confintTO_matched$upper,3)`  

To conclude: There is enough evidence to suggest the GOTV call increases the likelihood of voting  
in the 1998 election because the interval is positive and 0 is not included in the confidence interval.  